---
title: "Docker"
author:
- name: Pascal Maas
  affiliation: >
   Leiden Academic Centre for Drug Research, 
   Leiden University, Netherlands
  email: p.maas@lacdr.leidenuniv.nl
  
date: "`r Sys.Date()`"
output: 
    html_document:
        theme: united
        highlight: tango
    vignette: >
        %\VignetteIndexEntry{Docker}
        %\VignetteEngine{knitr::rmarkdown}
        \usepackage[utf8]{inputenc}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(devtools)
```

## Introduction
[Docker](https://docker.com) is software for virtualizing software in 
containers. These containers are packaged with necessary dependencies for the
software to run. This enables us to run any software anywhere. It is possible
to run `sumR` with Docker. It is assumed that Docker is installed
on your system.

## Dockerfile
The Dockerfile contains the instructions for building and running a contrainer.
We start by downloading a base image, `shiny-verse`. We then install the 
latest version of `sumR` from GitLab. The last line indicates what command
should be executed when we run the container. This file should be saved as 
`Dockerfile` without extensions.

```dockerfile
FROM rocker/shiny-verse:4.2.1

...

...
```

## Building container
After creating the Dockerfile, we can build the container. Open a terminal
in the folder where the Dockerfile is stored and run the following command to
build a container with the tag `sumR`. This process may take a few 
minutes to run, since dependencies need to be downloaded and installed.

```dockerfile
docker build -t sumR .
```

## Running Docker
Once the container has been built, we can run untargetedms by calling the
`docker run` command. 
```dockerfile
docker run sumR
```

# SessionInfo
```{r, echo = FALSE}

session_info("attached")
```
