---
title: "Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

_This package is under heavy development. Please do not use this package for actual reserach._

```{r setup}
library(sumR)
```

First we load up some example data. Here we have 4 injection of which 1 is a blank sample, all with negative polarity. These files have spectra which are combined together and centroided. 
```{r files}
dir <- system.file("extdata", package = "sumR")
files <- list.files(dir, pattern = ".*.mzML$", full.names = T)
files
```

We're using `xcms` in order to read our mzML files. 
```{r reading}
data <- read_mzml(files)
```

Next, we are identifying peaks and grouping them into features using the `dims_pipeline`. During this process, we also filter any chromatic peaks using an exclusion filter. In `feature_processing` we combine some of our feature data in order to make it ready for isotope detection.

```{r processing}
data <- dims_pipeline(data)
```

```{r features, fig.width=12}
data
df <- feature_processing(data)
df
```


Here we run the isotope detection algorithm. It adds several columns to our previous data.frame which can be used to filter isotopes from our feature set. 

```{r isotopes}
df <- isotope_tagging(df)
```
```{r normalize}
df2 <- normalize_features(df)

```

```{r pca, fig.width=12}

intensity_df <- df2[, grep("*.mzML$", colnames(df2))]
pca <- prcomp(intensity_df, center = TRUE, scale. = TRUE)
if ("ggbiplot" %in% installed.packages()) {
  ggbiplot::ggbiplot(pca)
} else {
  plot(pca$x)
}
```

```{r umap}
if ("umap" %in% installed.packages()) {
  library(umap)
  library(ggplot2)
  map <- umap(intensity_df)
  df = as.data.frame(map$layout)
  colnames(df) <- c("x", "y")
  maxes <- as.vector(apply(intensity_df, 1, which.max))
  col_vec <- c("LOW", "LOW", "UP", "UP", "UP", "UP")
  df$type <- col_vec[maxes]
  ggplot(df, aes(x = x, y = y, color = type)) + 
    geom_point() +
    ggtitle("Maximum intensity of each feature colored by group")
}

```

```{r sessionInfo}
sessionInfo()
```
